Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-05 discussion.tex
small improvement
cleaning up...
cleaning up till performance

>>>>

- detecting minor versions which don't need to be tracked, since they didn't touch the performance of the system. E.g. a modification of the README, etc.

- discuss later one situation in which opt-out might be desirable.

- tracking other types of information: exceptions, etc. especially since logs usually get lost... nobody has time to look in them...

### 2018-04-05 introductory.tex
cleaning up till performance

\section{Case Study: The Technology (Flask)}
\label{sec:tool}

@@

\subsection*{ENSURE CLARITY: NOTHING PREVENTS GENERALIZATION}
In this article we take as an example the Flask case study
and of the \tool as a dashboard. 
But there is nothing that prevents the generalization of 
our approach to other languages and technologies.

>>>>

\subsection*{Case Study: The Technology (Flask)}
\label{sec:tool}

@@

\subsection*{This is a LONGITUDINAL STUDY}

The dashboard has been deployed with the case stuy API for 
ten months, time in which more than 112.000 hits of the API have 
been recorded.

\subsection*{TO ARGUE: NOTHING PREVENTS GENERALIZATION}
In this article we take as an example the Flask case study
and of the \tool as a dashboard. 
But there is nothing that prevents the generalization of 
our approach to other languages and technologies.

### 2018-04-05 loc1.tex
cleaning up till performance

  The remainder of this section presents several interactive
  visualizations that are available in the dashboared without any further configuration\footnote{We recommend obtaining a color version of this paper for better readability}.

>>>>

  The remainder of this section presents several interactive
  visualizations that are available in the dashboared without any further configuration\footnote{We recommend obtaining a color version of this paper for better readability}. They are focused on two types of information that is important to the API maintainer:

  \begin{itemize}

    \item {\bf Utilization} -- refers to how the third parties use the API, which parts are most used, which are little used, etc. This is known to be critical information for upstream developer in general\cite{Haen14a}. In the case of source code dependencies one can detect such information by analyzing software repositories. However, in the case of services, there is no other way but monitoring service utilization.

    \item {\bf Performance} as measured in response times. This is important, especially for APIs that are supposed to be integrated in live systems.

  \end{itemize}

### 2018-04-05 utilization.tex
cleaning up...
cleaning up till performance

\newpage
\subsection{Utilization}
\label{sec:util}

\ins{Knowing how third parties use one's API is difficult even in the case of static dependencies. In the case of services, there is no other way but monitoring service utilization. \tool introduces a series of perspectives on utilization. }

\ins{

  The most basic possible view shows the cummulative information about all the endpoints and the number of calls to that endpoint over the lifetime of its tracking as well as in the current day and the last seven days.  
}

  \begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{utilization-table}
  \caption{....}
  \label{fig:basicest}
  \end{figure}

This kind of information is already very useful, and can provide the API maintainer with insight into the evolution of their system. By looking at \Fref{fig:basicest}, which presents the top 7 endpoints (for lack of space) one can already see several patterns:

@@

  \item {\bf Application Specific Patterns of System Usage}. The two most used endpoints in the figure stand for the two main activities in the system:

@@

\lesson{Although very useful, it was only after several months of using the system that the client requested the extra columsn for one day and seven days.}

  % \begin{figure}[h!]
  % \centering
  % \includegraphics[width=0.5\linewidth]{number_of_requests_}
  % \caption{The number of requests per endpoint per day view shows the overall utilization of the monitored application}
  % \label{fig:aeu}
  % \end{figure}

The table view is limited, and the one day and one week periods are conveniently but arbitrarily selected. For a more detailed evolution of utilization, Figure \ref{fig:aeu} shows the \perspective{Daily API Utilization} perspective on endpoint utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows a peak utilization, a day when the API had more than 12.000 hits.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{number_of_requests_}
\caption{Usage patterns become easy to spot in the requests per hour heatmap}
\label{fig:aeu}
\end{figure}

% \begin{figure}[h!]
%   \centering
%   \subfloat[Daily Utilization]{\includegraphics[width=.42\columnwidth]{number_of_requests_}\label{fig:aeu}}
%   \subfloat[Hourly Utilization]{\includegraphics[width=.5\columnwidth]{daily_patterns_}\label{fig:dp}}
%   \caption{Some of the available views\label{fig:views}}
% \end{figure}

% The way users interact with the platform can also be inferred since the endpoints are indicators of different activity types, e.g.:

% Besides showing the overall utilization, this endpoint provides the maintainer with information relevant for decisions regarding endpoint deprecation --- one of the most elementary ways of {\em understanding the needs of the downstream}\cite{Haen14a}. In our case study, the maintainer realized that one endpoint which they thought was not being used (i.e. \code{words\_to\_study}), contrary to their expectations, was actually being used\footnote{A complementary type of usage information can also be discovered in the view presented in Figure \ref{fig:sep} where seeing that an endpoint is never accessed can increase the confidence of the maintainer that a given endpoint is not used, although it can never be used a proof.}.

% \niceseparator

%   \todo{Add the time series graph and discuss it before the heatmap? We can then sell the heatmap better} 
%   \ml{Not sure about which graph you refer to here V}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{daily_patterns_}
\caption{Usage patterns become easy to spot in the requests per hour heatmap}
\label{fig:dp}
\end{figure}

>>>>

\subsection{Utilization}
\label{sec:util}

  The default landing page of the dashboard presents an overview of all the endpoints, 
  overview which presents some basic utilization information as:
  (1) the number of calls to that endpoint since the begining of tracking
  (2) the number of calls in the last seven days, and
  (3) the number of calls in the last day\footnote{Although very useful, it was only after several months of using the system that the client requested the extra columsn for one day and seven days. An argument for slow research}s

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\linewidth]{utilization-table}
      \caption{The API Overview presents synthetic endpoint utilization information}
      \label{fig:basicest}
    \end{figure}

  This kind of information is already very useful, and can provide the API maintainer with insight into the evolution of their service API. The seven endpoints present in \Fref{fig:basicest} illustrate three types of insight that the maintainer can obtain by looking at such a view:

@@

    \item {\bf Service Specific Usage Patterns}. The two most used endpoints in the figure stand for the two main activities in the system:

@@

\niceseparator

  For a more detailed evolution of utilization, Figure \ref{fig:aeu} shows the \perspective{Daily API Utilization} perspective on API utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows a peak utilization, on June 28 when the API had more than 12.000 hits\footnote{A highschool of students that were using the system for their class had a deadline on that date}.

    \begin{figure}[!ht]
    \centering
    \includegraphics[width=0.8\linewidth]{number_of_requests_}
    \caption{Usage patterns become easy to spot in the requests per hour heatmap}
    \label{fig:aeu}
    \end{figure}

\niceseparator

@@

    \begin{figure}[!ht]
      \centering
      \includegraphics[width=0.8\linewidth]{daily_patterns_}
      \caption{Usage patterns become easy to spot in the requests per hour heatmap}
      \label{fig:dp}
    \end{figure}
